---
title: "Networks of financial actors"
author: "Juan Rocha"
date: "Last update `r Sys.Date()`"

output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = TRUE, message=FALSE, warning = FALSE, error = FALSE,
    fig.width = 6, fig.height = 5, tidy = TRUE)
```

```{r message = FALSE, warning = FALSE, error = FALSE}
library(tidyverse)
library(plotly)
library(patchwork)
library(network)
library(sna)

theme_set(theme_light(base_size = 10))
```

## Read data

```{r message=FALSE, warning=FALSE, error=FALSE}
df1 <- readxl::read_xlsx(
    path = "data/Paulas_files/001-Drivers_Increasing_Risk_ED.xlsx",
    sheet = 2,
    na = "NA") %>% 
    janitor::clean_names()

```

Currently the data set has `r df1 %>% nrow()` companies, in `r df1$country %>% unique() %>% length()` countries, dealing with `r df1$commodity %>% unique() %>% length()` commodities (including `NA`). Below some overview of the missing values:

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    skimr::skim()
```

## Networks

### Countries 

Connected by companies and resource flow

```{r}

net_countries <-  df1 %>% 
  select(company, origin, country) %>% 
  group_by(origin, country) %>% 
  # unique() %>% 
  count(name = "channel") %>% # a channel is a country * commodities
  filter(!is.na(origin), !is.na(country)) %>% 
  network(directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", ignore.eval=FALSE, loops = TRUE)

net_countries %v% "indegree" <- sna::degree(
  net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
  cmode = "indegree")

net_countries %v% "outdegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
    cmode = "outdegree")

net_countries %v% "degree" <- sna::degree(
  net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE)

gplot(
    net_countries,
    vertex.col = alpha("purple", 0.25),
    label = network.vertex.names(net_countries),
    label.cex = 0.75, label.col = "grey50",
    vertex.border = 0, 
    vertex.cex = sqrt(0.2 + net_countries %v% "outdegree"),
    label.pos = 5,
    edge.col = alpha("grey50", 0.1),
    edge.lwd = log1p(net_countries %e% "channel"),
    displayisolates = TRUE, pad = 0.5,
    mode = "mds"
    )


```

```{r include = FALSE}
# Alternative visualizations:
# networkD3::simpleNetwork(
#     df1 %>%
#         select(company, origin, country) %>%
#         group_by(origin, country) %>%
#         # unique() %>%
#         count(name = "channel") %>% 
#         filter(!is.na(origin), !is.na(country)) ,
#     Source = "origin", Target = "country", zoom = TRUE,
#     linkColour = "grey", nodeColour = "blue")
library(ggnetwork)

g1 <- ggplot(ggnetwork(net_countries, arrow.gap = 0.01), 
       aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(
        aes(color = channel), size = 0.15, alpha = .5,
        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = outdegree), color = "red", alpha = 0.4) +
    geom_nodes(aes(size = indegree), color = "blue", alpha = 0.4) +
    geom_nodetext(aes(label = vertex.names, size = degree/4)) +
    scico::scale_color_scico(
        "Channel", palette = "batlow", direction = -1,
        guide = guide_colorbar(
            barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
    scale_size(name = "Degree", range = c(1,2)) +
    labs(title = "Network of countries connected by companies",
         subtitle = "A channel is a company times commodities. Red nodes are influencers while blue are influenced per\nnumber of connections", caption = "Data source: Orbis database") +
    theme_blank(base_size = 6) 
# 
# ggsave(
#     filename = "figures/countries_net.png",
#     plot = g1,
#     device = "png",
#     width = 4, height = 3,
#     bg = "white"
# )

```


Maybe fit a PCA to the data with network stats, commodities and so forth per country, and use the coordinates for visualization. Currently using a multi-dimensional scaling projection. Another way of "seeing" the network is by ranking which countries are the most influential given their links to other countries (_outdegree_), once self-links are removed (companies whose country of origin and operation is the same):

```{r}
df1 %>% 
    select(company, origin, country) %>% 
    group_by(origin, country) %>% 
    # unique() %>% 
    count(name = "channel") %>% # a channel is a country * commodities
    filter(!is.na(origin), !is.na(country)) %>% 
    filter(origin != country) %>% 
    arrange(desc(channel))
```

The dataset has `r unique(df1$company) %>% length()` unique companies and `r unique(df1$commodity) %>% length()` commodities. Below a table that summarizes in how many different places the same company extract the same resource:

```{r}
df1 %>% 
    select(company, commodity) %>% 
    group_by(company, commodity) %>% 
    count(name = "geographies") %>% 
    arrange(desc(geographies))
```



```{r fig.width = 6, fig.height = 5}
g2 <- df1 %>% 
  select(company, commodity) %>% 
  group_by(company, commodity) %>% 
  count(name = "geographies") %>% 
  arrange(desc(geographies)) %>% 
  filter(geographies > 3) %>%  ungroup() %>% 
  mutate(company = as_factor(company)) %>%
  ## the data needs to be ungrouped for fct_reorder to work!
  mutate(company = fct_reorder(
    .f = company, .x = geographies, .fun = sum, .desc = FALSE)) %>% 
  #pull(company) %>% levels()
  ggplot(aes(y = company)) +
  geom_bar(aes(weight = geographies, fill = commodity),
           position = "stack", show.legend = TRUE, alpha = 0.5) +
  scale_fill_discrete(
    "Commodities", guide = guide_legend(
      ncol = 4, title.position = "top", title.hjust = 0.5,
      keywidth = 0.5, keyheight = 0.5
    )
  ) +
  labs(title = "Companies ranking by number of countries and commodities",
       subtitle = "A geography is a combination of `country * commodity` as a proxy of geographical influence of a company.\nOnly companies with >3 geographies are shown", caption = "Data curated by the project, source Orbis database", x = "Geographies", y = "Top companies") +
  theme_light(base_size = 5) + theme(legend.position = c(0.7, 0.1))


# ggsave(
#     filename = "figures/companies_commodity.png",
#     plot = g2,
#     device = "png",
#     width = 6, height = 5,
#     bg = "white"
# )

g2
```

Note there is a relatively low diversity of specialization within companies, meaning that a lot of them are putting their eggs on the same basket. This is not surprising. The knowledge and know-how of a company is limited and can only produce a set of products that are themselves similar in terms of the tech, knowledge and institutions.

### Companies 

Companies connected by shared countries*resources.

Another key network representation is by companies given the countries where they together operate, and by the common geographies (meaning exporting the same commodity from the same country).

```{r}
bip_net <- df1 %>% 
  select(company, origin, country) %>% 
  # Assumption: if country is NA, it means the company operates in
  #  its country of origin:
  mutate(country = case_when(
    is.na(country) ~ origin, !is.na(country) ~ country))  %>% 
  group_by(company, country) %>% 
  # a channel can be either different commodities, or different places within the same country
  count(name = "channel") %>% 
  ungroup() |> 
  network(
    directed = TRUE, bipartite = TRUE, matrix.type = "edgelist", 
    ignore.eval=FALSE, loops = FALSE)
 
bip_mat <- bip_net |> 
  as.matrix.network.adjacency()
```

The bipartite matrix has `r dim(bip_mat)[1]` companies and `r dim(bip_mat)[2]` countries.

```{r}
net_comp <- network(
  bip_mat %*% t(bip_mat), 
  directed = FALSE)

net_comp %v% "degree" <- sna::degree(
  net_comp, gmode = "graph", rescale = FALSE)
net_comp %v% "betw" <- sna::betweenness(
  net_comp, gmode = "graph", rescale = TRUE)

df_comps <- tibble(
  companies = network.vertex.names(net_comp),
  degree = net_comp %v% "degree",
  betweenness = net_comp %v% "betw"
)

df_comps |> 
  arrange(desc(degree))
```

```{r fig.cap="Network of companies working on the same countries regardless of commodity. Isolates has been excluded"}
plot.network(
  net_comp,
  edge.col = alpha("grey50", 0.3),
  edge.lwd = 0.1,
  vertex.lty = 0,
  vertex.cex = 0.2,
  arrowhead.cex = 0.3,
  displayisolates = FALSE)

g3 <- ggplot(ggnetwork(net_comp, arrow.gap = 0.01),
    aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(color = "grey40", size = 0.15, alpha = .5,
        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = degree), color = "red", alpha = 0.4) +
    geom_nodes(aes(size = betw), color = "blue", alpha = 0.4) +
    #geom_nodetext(aes(label = vertex.names, size = degree/4)) +
    scico::scale_color_scico(
        "Channel", palette = "batlow", direction = -1,
        guide = guide_colorbar(
            barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
    scale_size(name = "Degree", range = c(1,5)) +
    labs(title = "Network of companies",
         subtitle = "Companies are connected if they share geographies", caption = "Data source: Orbis database") +
    theme_blank(base_size = 6) 



# ggsave(
#     filename = "figures/companies_net.png",
#     plot = g3,
#     device = "png",
#     width = 4, height = 3,
#     bg = "white"
# )
```


```{r}
bip_net <- df1 %>% 
  select(company, origin, country) %>% 
  # Assumption: if country is NA, it means the company operates in
  #  its country of origin:
  mutate(country = case_when(
    is.na(country) ~ origin, !is.na(country) ~ country))  %>% 
  group_by(company, country) %>% 
  # a channel can be either different commodities, or different places within the same country
  count(name = "channel") %>% 
  ungroup() |> 
  network(
    directed = TRUE, bipartite = TRUE, matrix.type = "edgelist", 
    ignore.eval=FALSE, loops = FALSE)
 
bip_mat <- bip_net |> 
  as.matrix.network.adjacency()
```

### Financial actors

Investors linked by investments on the same companies:

```{r}
fls <- list.files(path = "data/Paulas_files/Financial_actors/")
fls <- paste0("data/Paulas_files/Financial_actors/", fls)
head(fls)
```

```{r financial-actors}
dat <- map(
  fls,
  function(x) {
    x |> 
      read_delim(
        delim = ";", col_types = "cccccccccc", name_repair = "minimal",
        col_names = c("company", "name", "pas", "type", "direct_percent", "total_percent", "source", "date", "op_revenue", "employees"),
        skip = 1, na = c("n.d.", "-")) |> 
    janitor::clean_names() }
) 


dat <- dat |> 
  bind_rows() 
  
dat <- dat |> 
  filter(company != "") |> # 2 empty rows
  mutate(
    company = str_to_title(company),
    name = str_to_title(name),
    total_percent = str_remove_all(total_percent, pattern = "±")
  ) |> 
  mutate(
    total_percent = as.numeric(total_percent),
    direct_percent = as.numeric(direct_percent), 
    date = lubridate::my(date),
    op_revenue = as.numeric(op_revenue),
    employees = as.numeric(employees)
  )
  
  
  
skimr::skim(dat)

# save(dat, file = "data/investors_cleaned.RData")
```

The dataset contains `r dat |> pull(name) |> unique() |> length()` investors and `r dat |> pull(company) |> unique() |> length()` companies extracting resources in areas prone to the emergence of zoonotic dieseases. 

```{r}
# note there are companies that are investors in others, so it's not purely bipartite
net_companies <- dat |> 
  select(investors = name, company, total_percent) |> 
  network(
    directed = TRUE, bipartite = FALSE, matrix.type = "edgelist", 
    ignore.eval=FALSE, loops = FALSE)

net_companies %v% "outdegree" <- sna::degree(
    net_countries, gmode = "digraph", diag = TRUE, rescale = FALSE, 
    cmode = "outdegree")
```

```{r}
plot.network(
  net_companies,
  edge.col = alpha("grey50", 0.3),
  edge.lwd = 0.1,
  vertex.lty = 0,
  vertex.cex = 0.2,
  arrowhead.cex = 0.3
  )

g4 <- ggplot(ggnetwork(net_companies, arrow.gap = 0.01),
    aes(x = x, y = y, xend = xend, yend = yend)) +
    geom_edges(color = "grey40", size = 0.15, alpha = .5,
        arrow = arrow(length = unit(2, "pt"), type = "closed")) +
    geom_nodes(aes(size = outdegree), color = "orange", alpha = 0.4) +
    #geom_nodes(aes(size = betw), color = "blue", alpha = 0.4) +
    #geom_nodetext(aes(label = vertex.names, size = degree/4)) +
    # scico::scale_color_scico(
    #     "Channel", palette = "batlow", direction = -1,
    #     guide = guide_colorbar(
    #         barwidth = unit(3, "mm"), barheight = unit(1.5, "cm"))) +
    scale_size(name = "Outdegree", range = c(0.2,4)) +
    labs(title = "Network of shareholders",
         subtitle = "Shareholders are connected if they have shares on the same companies", caption = "Data source: Orbis database") +
    theme_blank(base_size = 6) 

# ggsave(
#     filename = "figures/shareholders_net.png",
#     plot = g4,
#     device = "png",
#     width = 4, height = 3,
#     bg = "white"
# )
```

```{r}
out_df <- tibble(
  actor = network.vertex.names(net_companies),
  outdegree = net_companies %v% "outdegree"
)

out_df |> 
  ggplot(aes(outdegree)) +
  geom_density()
```

There is a few investors with a disproportionate number of investments in the network.

```{r}
out_df |> 
  arrange(desc(outdegree))
```

